Consistent Polyhedral Surrogates via Embedding

### 2021-07-28 notes/top-k.tex
Adding in summary from meeting notes 27 July

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and narrow down to a finite representative set for $L^{top-k}$.

\begin{claim}
	For each $p \in \mathrm{relint}\simplex$, the expected loss $\inprod{p}{L^{top-k}(u)}$ is affine on the set $\{u \in \U \mid $ the top-$k$ elements of $(u - e_y)$ are the same for all $ y \in \Y\}$.
\end{claim}
\begin{proof}
	For $u \in \U$, the loss for any outcome is given by $L^{top-k}(u)_y = \left(1 - u_y + \frac 1 k \sum_{i = 1}^k(u - e_y)_{[i]}\right)$.
	Since this is affine in the first two terms, it is thus affine on any set for which the last sum is composed of the same elements, which is precisely given by the set above.
\end{proof}

\jessie{This is not precise...}
\begin{proposition}
	$\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}$ is the set of vertices of $L^{top-k}$.
\end{proposition}
\begin{proof}
	Any vertex of the loss surface in $\reals^n \cap \U$ is given by the intersection of (at least) $n-1$ affine halfspaces.
	The set of affine halfspaces composing $L^{top-k}$ is precisely those where the top-$k$ elements of $(u - e_y)$ are the same for all $y \in \Y$.
	These sets shift when either (a) one of the $k$ nonzero elements shifts below $1$ or (b) we hit the boundary of $\U$.
	For (a), this is the intersection of the $n$ hyperplanes where \jessie{...}
	When we hit the boundary of $\U$, the vertices are \jessie{...}
\end{proof}

>>>>

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and narrow down to a finite representative set $\R$ for $L^{top-k}$.
Consider $\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}$.

\begin{claim}
	For each $p \in \mathrm{relint}\simplex$, the vertices of the expected loss $\inprod{p}{L^{top-k}(u)} = 1 - \inprod{p}{u} + \sum_{i=1}^k u_{[i]} - \inprod{p}{\min(u,\ones)}$ are denoted by $\R$.
\end{claim}
\begin{proof}
	First, observe the loss is affine in the first two terms for all $u \in \U$, and non-linearity is only introduced in the last two terms.
	Without loss of generality, we will fix a partition $\pi$ of the elements of $u$, and let $T_k(u)$ be the top $k$ elements of $u$ according to the partition $\pi$.
	Since $u \in \U \supseteq R^{\tt low}$, we have $u_{\pi_{k+1}} = \ldots = u_{\pi_n} = 0$, and take $k$ constraints constructing $R^{\tt low}$ given by $0 \leq u_{\pi_i}$ for $i \in [k]$.
	Call $T$ the top $k$ elements of the partition $\pi$.

	Moreover, there are $k$ constraints constructing $R^{\tt high}$, given by $	\forall i \in T_k(u), \,\,
	u_{i} \leq 1 + \frac{1}{k-1}\sum_{j \in T_k(u) \setminus \{i\}} u_j$.

	%need to introduce the segments where L is affine.
	Within each permutation $\pi$, we know that the third term is affine on each set $\{u \in \U \mid $ the top $k$ elements of $u$ are equal to $T\}$.

	Finally, we can consider affineness in the last term, where there is a ``switch'' at $u_y = 1$ for each $y$.
	There are $2^k$ choices of sets $V \in \{A,B\}^k$, $u \in V$ if and only if, for all $y \in \Y$, $V_y = A$ means $u_y \leq 1$ and $V_y = B$ otherwise. 
	We have $L$ affine on the region $\{u \in \U \mid u \in V\}$ for each choice of $V \in \{A,B\}^k$.
	We will generate such sets using $T$ and $S$, the latter of which is introduced below.

	%Now, since each of these inequalities generates a halfspace in $\reals^k$, and a vertex in $\reals^k$ is the intersection of $k$ such halfspaces, we observe that there are $2^k$ possible vertices (though some of these are redundant).

	For a fixed $T$ (the top $k$ elements of $\pi$), iterate over every $S \subseteq T$.
	Suppose $S$ is the set of elements of $u$ such that $0 \leq u_y \leq 1$ for all $y \in S$, and additionally have the constraints $1 \leq u_y \leq 1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j$.
	For each $T,S$ combination, we then have vertices at each of these $2^k$ possible equalities.
	\begin{align*}
	\forall i \in S, \,\, 0 &\leq u_i \leq 1 \\
	\forall i \in T \setminus S, \,\, 1 &\leq u_i \leq 1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j 
	\end{align*}
	Iterating over each of these $2^k$ inequalities, we see that vertices are generated at points $0,1$ or some constant $c_{TS}$ for each choice of inequalities for $T$ and $S$.

	Now, for each $T$ and $S$ and choice of equalities in each, let $M = \{y \mid u_y = 1\}$, and $H = \{y \in \Y \mid u_y = 1 + \sum_{j \in T_k(u) \setminus \{i\}} u_j\}$.
	(Note that $y \in H \implies u_y > 1$ unless $M$ is empty \jessie{and $H = \{y\}$?}, in which case, we will say $y \in M$ and $H = \emptyset$.) 
	\jessie{Be careful about if $c_{MH} = 1$}
	By construction of $T$ and $S$, we satisfy the set requirements of $\R$, and observe that for all $y \in H$, we have $u_y = 1 + \frac 1 {k-1}\sum_{j \in T_k(u) \setminus \{y\} u_j} = \frac{|M| + k -1}{k - |H|}$.

	As there are only finitely many choices of $T$ and $S$, a finite number of vertices generated by each of those choices, and they generate the sets generating $\R$.
	Moreover, for each $H,M$ generating an element of $R$, there is a $T,S$ generating the element of $\R$.
	Thus, we conclude $\R$ is the finite set of vertices of $\inprod{p}{L(\cdot)}$ for distributions $p \in \mathrm{relint}(\simplex)$.
\end{proof}

%\jessie{This is not precise...}
%\begin{proposition}
%	is the set of vertices of $L^{top-k}$.
%\end{proposition}
%\begin{proof}
%	Any vertex of the loss surface in $\reals^n \cap \U$ is given by the intersection of (at least) $n-1$ affine halfspaces.
%	The set of affine halfspaces composing $L^{top-k}$ is precisely those where the top-$k$ elements of $(u - e_y)$ are the same for all $y \in \Y$.
%	These sets shift when either (a) one of the $k$ nonzero elements shifts below $1$ or (b) we hit the boundary of $\U$.
%	For (a), this is the intersection of the $n$ hyperplanes where \jessie{...}
%	When we hit the boundary of $\U$, the vertices are \jessie{...}
%\end{proof}
